---
title: "Simulation of Exponential Distribution"
author: "Anthony Cato"
date: "April 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
```

### Overview
In this project, I am investigating how the exponential distribution compares with the Central Limit Theorem. I will create 1000 simulations to observe this distribution and leverage the base plotting package to view this exponential distribution. 

### Simulation 
```{r}
# Default lambda value for investigation will be 0.2
lambda <- 0.2
# Number of exponentials used for each simulation
n <- 40
# Number of simulations
B <- 1000

# Setting the seed to make project reproducible
set.seed(7919)

# Generates 40000 random exponential values
exponentials <- rexp(n * B, lambda)

# Creates matrix of exponential values
simulations <- matrix(exponentials, nrow = B)

# Converts matrix to a data.table and computes the mean for each column
means <- data.table(apply(simulations, 1, mean))
```
### Sample Mean versus Theoretical Mean
```{r}
theoreticalMean <- 1 / lambda 
sampleMean <- mean(means$V1)

print(c(theoreticalMean, sampleMean))
```
### Sample Variance versus Theoretical Variance
```{r}
theoreticalVariance <- (1 / (lambda ^ 2)) / n 
sampleVariance <- var(means$V1)

print(c(theoreticalVariance, sampleVariance))
```
### Distributions
```{r}
# Plotting the distribution of the means
hist(means$V1)
# Drawing a line at x = sampleMean
abline(v = sampleMean, col = "green")
```

### Conclusion
A cursory look at the exponential distribution reveals that the average values are centered around the mean and resembles the normal distribution.